---
title: "PLSDA"
author: "Alex Muñoz"
date: "6/5/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PLS-DA analysis 

We have performed PLS-DA over the results provided by k-means in order to determine which of the attributes were the most influential on every cluster and reduce the number of attributes for showing the profile of the patient that falls in each cluster.
```{r}
library(openxlsx)
library(dplyr)
data_ps = read.xlsx("C:/CIENCIA DE DATOS/UPV 3º CURSO/2º SEMESTRE/PROY III/Kmedoids_melanoma.xlsx")
names(data_ps)
data_ps.def = data_ps %>% dplyr::select(-1,-41:-56)
grupo_etio = data_ps$Grupo_etiopatogénico
Y = as.factor(data_ps$kmed_clus)
data_ps.def = data_ps.def %>% dplyr::select(-32,-40)
```

```{r}
cat_vars = data_ps.def %>% dplyr::select(-2:-5,-10:-11,-32,-35)
data_ps.def[colnames(cat_vars)] <- lapply(data_ps.def[,colnames(cat_vars)], factor)
```
We have created dummies for the factor features to join with the numeric original variables and introduce them as input for the PLS-DA model. 

```{r}
library(fastDummies)
library(scales)
library(kableExtra)
con_vars = data_ps.def %>% dplyr::select(2:5,10:11,32,35)
con_vars$Edad = rescale(con_vars$Edad)
con_vars$Peso = rescale(con_vars$Peso)
con_vars$Altura = rescale(con_vars$Altura)
con_vars$Añosprofsol = rescale(con_vars$Añosprofsol)
con_vars$IMC = rescale(con_vars$IMC)
con_vars$Añospaquete = rescale(con_vars$Añospaquete)
con_vars$Breslow = rescale(con_vars$Breslow)
con_vars$Mitosis = rescale(con_vars$Mitosis)
cat_vars = data_ps.def %>% dplyr::select(-2:-5,-10:-11,-32,-35)
dummies = dummy_cols(cat_vars)
dummies = dummies %>% dplyr::select(-1:-30)
data_ps_dum = cbind(con_vars,dummies)
```
We set 10 initially as the number of components as in the package guide setting a number of cp's higher than the number of classes of the response variable is recommended.

```{r}
library(devtools)
library(mixOmics)
srbct.plsda <- plsda(data_ps_dum, Y, ncomp = 10)
```
The perf evaluates the performance of PLS-DA for a large number of components, using repeated k-fold cross-validation

```{r} 
perf.plsda.srbct <- perf(srbct.plsda, validation = "Mfold", folds = 3, 
                  progressBar = FALSE, auc = TRUE, nrepeat = 10,cpus = 2)
```
```{r}
perf.plsda.srbct$choice.ncomp
plot(perf.plsda.srbct, col = color.mixo(5:7), sd = TRUE, legend.position = "horizontal")
```

The plot outputs the classification error rate, or Balanced classification error rate when the number of samples per group is unbalanced, the standard deviation according to three prediction distances. Here we can see that for the BER and the Mahalannobis distance, the best performance seems to be achieved for ncomp = 3.  

### Tuning parameters
```{r}
tune.splsda.srbct <- tune.splsda(data_ps_dum, Y, ncomp = 9, validation = 'Mfold', folds = 5, 
                           progressBar = TRUE, dist = 'mahalanobis.dist', measure = "BER",
                           nrepeat = 10, cpus = 2)
```

```{r}
ncomp <- tune.splsda.srbct$choice.ncomp$ncomp # optimal number of components based on t-tests
ncomp
```
We can then extract the classification error rate averaged across all folds and repeats for each tested keepX value, the optimal number of components, the optimal number of variables to select per component which is summarized in a plot where the diamond indicated the optimal keepX value:

```{r}
select.keepX <- tune.splsda.srbct$choice.keepX[1:ncomp]  # optimal number of variables to select
select.keepX
plot(tune.splsda.srbct, col = color.jet(9))
```
Based on those tuning results, we can run our final and tuned sPLS-DA model:

```{r}
splsda.srbct <- splsda(data_ps_dum, Y, ncomp = 9, keepX = select.keepX)
loadings = data.frame(abs(splsda.srbct[["loadings"]][["X"]]))
```
```{r}
plotIndiv(splsda.srbct, comp = c(1,2),
          group = data_ps$Grupo_etiopatogénico, ind.names = FALSE, legend = TRUE,
          title = 'sPLS-DA on SRBCT, comp 1 & 2')
```

```{r}
compo1 =  data.frame(loadings[which(loadings$comp1 != 0),])
c1_load = ggplot(compo1, aes(x = rownames(compo1), y = comp1,fill=comp1,group=comp1)) + geom_col() +scale_fill_gradient(high = "green", low = "palegreen")+
  labs(title = "Loadings Comp 1",x = NULL, y = "Weights of features")+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+theme(plot.title = element_text(hjust = 0.5))+theme(legend.position = "none")
```
```{r}
compo2 =  data.frame(loadings[which(loadings$comp2 != 0),])
c2_load = ggplot(compo2, aes(x = rownames(compo2), y = comp2,fill=comp2,group=comp2)) + geom_col() +scale_fill_gradient(high = "green", low = "palegreen")+
  labs(title = "Loadings Comp 2",x = NULL, y = "Weights of features")+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+theme(plot.title = element_text(hjust = 0.5))+theme(legend.position = "none")
```
```{r}
compo3 =  data.frame(loadings[which(loadings$comp3 != 0),])
c3_load = ggplot(compo3, aes(x = rownames(compo3), y = comp3,fill=comp3,group=comp3)) + geom_col() +scale_fill_gradient(high = "green", low = "palegreen")+
  labs(title = "Loadings Comp 3",x = NULL, y = "Weights of features")+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+theme(plot.title = element_text(hjust = 0.5))+theme(legend.position = "none")
```
```{r}
compo4 =  data.frame(loadings[which(loadings$comp4 != 0),])
c4_load = ggplot(compo4, aes(x = rownames(compo4), y = comp4,fill=comp4,group=comp4)) + geom_col() +scale_fill_gradient(high = "green", low = "palegreen")+
  labs(title = "Loadings Comp 4",x = NULL, y = "Weights of features")+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+theme(plot.title = element_text(hjust = 0.5))+theme(legend.position = "none")
```
```{r}
compo5 =  data.frame(loadings[which(loadings$comp5 != 0),])
c5_load = ggplot(compo5, aes(x = rownames(compo5), y = comp5,fill=comp5,group=comp5)) + geom_col() +scale_fill_gradient(high = "green", low = "palegreen")+
  labs(title = "Loadings Comp 5",x = NULL, y = "Weights of features")+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+theme(plot.title = element_text(hjust = 0.5))+theme(legend.position = "none")
```
```{r}
compo6 =  data.frame(loadings[which(loadings$comp6 != 0),])
c6_load = ggplot(compo6, aes(x = rownames(compo6), y = comp6,fill=comp6,group=comp6)) + geom_col() +scale_fill_gradient(high = "green", low = "palegreen")+
  labs(title = "Loadings Comp 6",x = NULL, y = "Weights of features")+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+theme(plot.title = element_text(hjust = 0.5))+theme(legend.position = "none")
```
```{r}
compo7 =  data.frame(loadings[which(loadings$comp7 != 0),])
c7_load = ggplot(compo7, aes(x = rownames(compo7), y = comp7,fill=comp7,group=comp7)) + geom_col() +scale_fill_gradient(high = "green", low = "palegreen")+
  labs(title = "Loadings Comp 7",x = NULL, y = "Weights of features")+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+theme(plot.title = element_text(hjust = 0.5))+theme(legend.position = "none")
```
```{r}
compo8 =  data.frame(loadings[which(loadings$comp8 != 0),])
c8_load = ggplot(compo8, aes(x = rownames(compo8), y = comp8,fill=comp8,group=comp8)) + geom_col() +scale_fill_gradient(high = "green", low = "palegreen")+
  labs(title = "Loadings Comp 8",x = NULL, y = "Weights of features")+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+theme(plot.title = element_text(hjust = 0.5))+theme(legend.position = "none")
```
```{r}
compo9 =  data.frame(loadings[which(loadings$comp9 != 0),])
c9_load = ggplot(compo9, aes(x = rownames(compo9), y = comp9,fill=comp9,group=comp9)) + geom_col() +scale_fill_gradient(high = "green", low = "palegreen")+
  labs(title = "Loadings Comp 9",x = NULL, y = "Weights of features")+theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+theme(plot.title = element_text(hjust = 0.5))+theme(legend.position = "none")
```
```{r}
library(cowplot)
plot_grid(c1_load, c2_load, c3_load, ncol = 3, nrow = 1)
plot_grid(c4_load, c5_load, c6_load, ncol = 3, nrow = 1)
plot_grid(c7_load, c8_load, c9_load, ncol = 3, nrow = 1)
```

```{r}
selected_feat = dplyr::select(data_ps,"Pelo", "MC1Rvar", "Efélides","Fototipo", "Ojos", "Qactínicas","Fotolocalización","TipoHX","Quemaduras_areaMM","CancerFam","Localización","Sexo","Quemaduras","Léntigos","Léntigos_en_área_de_MM","kmed_clus","Qactínicas","Nevuspre")
clus.1 = selected_feat %>%filter(kmed_clus==1) 
clus.1$kmed_clus=NULL
clus.2 = selected_feat %>% filter(kmed_clus==2)
clus.2$kmed_clus=NULL
clus.3 = selected_feat %>% filter(kmed_clus==3)
clus.3$kmed_clus=NULL
clus.4 = selected_feat %>%filter(kmed_clus==4)
clus.4$kmed_clus=NULL
clus.5 = selected_feat %>% filter(kmed_clus==5)
clus.5$kmed_clus=NULL
clus.6 = selected_feat %>% filter(kmed_clus==6)
clus.6$kmed_clus=NULL
clus.7 = selected_feat %>% filter(kmed_clus==7)
clus.7$kmed_clus=NULL
clus.8 = selected_feat %>% filter(kmed_clus==8)
clus.8$kmed_clus=NULL
```






## PATIENTS PROFILES

```{r}
library(survival)
library(openxlsx)
library(Hmisc)
library(ggplot2)
library(dplyr)
library(stringr)
library(FactoMineR)
library(factoextra)


data_ps = read.xlsx("Kmedoids_melanoma.xlsx")
```
```{r}

#separamos por cluster
selected_feat = dplyr::select(data_ps,"Pelo", "MC1Rvar", "Efélides","Fototipo", "Ojos", "Qactínicas","Fotolocalización","TipoHX","Quemaduras_areaMM","CancerFam","Localización","Sexo","Quemaduras","Léntigos","Léntigos_en_área_de_MM","kmed_clus","Nevuspre")
clus.1 = selected_feat %>%filter(kmed_clus==1)
clus.1$kmed_clus=NULL
clus.2 = selected_feat %>% filter(kmed_clus==2)
clus.2$kmed_clus=NULL
clus.3 = selected_feat %>% filter(kmed_clus==3)
clus.3$kmed_clus=NULL
clus.4 = selected_feat %>%filter(kmed_clus==4)
clus.4$kmed_clus=NULL
clus.5 = selected_feat %>% filter(kmed_clus==5)
clus.5$kmed_clus=NULL
clus.6 = selected_feat %>% filter(kmed_clus==6)
clus.6$kmed_clus=NULL
clus.7 = selected_feat %>% filter(kmed_clus==7)
clus.7$kmed_clus=NULL
clus.8 = selected_feat %>% filter(kmed_clus==8)
clus.8$kmed_clus=NULL
```


```{r}
# creamos dummies
library(fastDummies)
library(dplyr)
clus.1.dummies = dummy_cols(clus.1)
clus.1.dummies = clus.1.dummies %>% dplyr::select(-1:-16)

clus.2.dummies = dummy_cols(clus.2)
clus.2.dummies = clus.2.dummies %>% dplyr::select(-1:-16)

clus.3.dummies = dummy_cols(clus.3)
clus.3.dummies = clus.3.dummies %>% dplyr::select(-1:-16)

clus.4.dummies = dummy_cols(clus.4)
clus.4.dummies = clus.4.dummies %>% dplyr::select(-1:-16)

clus.5.dummies = dummy_cols(clus.5)
clus.5.dummies = clus.5.dummies %>% dplyr::select(-1:-16)

clus.6.dummies = dummy_cols(clus.6)
clus.6.dummies = clus.6.dummies %>% dplyr::select(-1:-16)

clus.7.dummies = dummy_cols(clus.7)
clus.7.dummies = clus.7.dummies %>% dplyr::select(-1:-16)

clus.8.dummies = dummy_cols(clus.8)
clus.8.dummies = clus.8.dummies %>% dplyr::select(-1:-16)
```


```{r}
datos1 = data_frame()
for(i in 1:38){
  a = c(sum(clus.1.dummies[i]), colnames(clus.1.dummies[i]))
  datos1 = rbind(datos1,a)
}

colnames(datos1) = c('valor', 'nombre')
datos1$valor = rescale(as.numeric(datos1$valor))

datos2 = data_frame()
for(i in 1:38){
  a = c(sum(clus.2.dummies[i]), colnames(clus.2.dummies[i]))
  datos2 = rbind(datos2,a)
}

colnames(datos2) = c('valor', 'nombre')
datos2$valor = rescale(as.numeric(datos2$valor))

datos3 = data_frame()
for(i in 1:39){
  a = c(sum(clus.3.dummies[i]), colnames(clus.3.dummies[i]))
  datos3 = rbind(datos3,a)
}

colnames(datos3) = c('valor', 'nombre')
datos3$valor = rescale(as.numeric(datos3$valor))

datos4 = data_frame()
for(i in 1:40){
  a = c(sum(clus.4.dummies[i]), colnames(clus.4.dummies[i]))
  datos4 = rbind(datos4,a)
}

colnames(datos4) = c('valor', 'nombre')
datos4$valor = rescale(as.numeric(datos4$valor))

datos5 = data_frame()
for(i in 1:39){
  a = c(sum(clus.5.dummies[i]), colnames(clus.5.dummies[i]))
  datos5 = rbind(datos5,a)
}

colnames(datos5) = c('valor', 'nombre')
datos5$valor = rescale(as.numeric(datos5$valor))


datos6 = data_frame()
for(i in 1:42){
  a = c(sum(clus.6.dummies[i]), colnames(clus.6.dummies[i]))
  datos6 = rbind(datos6,a)
}

colnames(datos6) = c('valor', 'nombre')
datos6$valor = rescale(as.numeric(datos6$valor))


datos7 = data_frame()
for(i in 1:40){
  a = c(sum(clus.7.dummies[i]), colnames(clus.7.dummies[i]))
  datos7 = rbind(datos7,a)
}

colnames(datos7) = c('valor', 'nombre')
datos7$valor = rescale(as.numeric(datos7$valor))

datos8 = data_frame()
for(i in 1:40){
  a = c(sum(clus.8.dummies[i]), colnames(clus.8.dummies[i]))
  datos8 = rbind(datos8,a)
}

colnames(datos8) = c('valor', 'nombre')
datos8$valor = rescale(as.numeric(datos8$valor))
```


```{r}
library(ggplot2)
library(ggpubr)
p1<-ggplot(datos1, aes(nombre,valor,fill=valor,group=valor)) + geom_col(width = 1) +theme(axis.text.x = element_text(size = 7,face="bold",angle = 90, vjust = 0.5, hjust=1))+scale_y_continuous(limits=c(0, 1))+scale_fill_gradient(high = munsell::mnsl("5P 2/12"), low = munsell::mnsl("5P 7/12"))+ggtitle("Cluster 1 Profile")+xlab(NULL)+ylab(NULL)+theme(legend.position = "none")

p2<-ggplot(datos2, aes(nombre,valor,fill=valor,group=valor)) + geom_col(width = 1) +theme(axis.text.x = element_text(size = 7,face="bold",angle = 90, vjust = 0.5, hjust=1))+scale_y_continuous(limits=c(0, 1))+scale_fill_gradient(high = munsell::mnsl("5P 2/12"), low = munsell::mnsl("5P 7/12"))+ggtitle("Cluster 2 Profile")+xlab(NULL)+ylab(NULL)+theme(legend.position = "none")

p3<-ggplot(datos3, aes(nombre,valor,fill=valor,group=valor)) + geom_col(width = 1) +theme(axis.text.x = element_text(size = 7,face="bold",angle = 90, vjust = 0.5, hjust=1))+scale_y_continuous(limits=c(0, 1))+scale_fill_gradient(high = munsell::mnsl("5P 2/12"), low = munsell::mnsl("5P 7/12"))+ggtitle("Cluster 3 Profile")+xlab(NULL)+ylab(NULL)+theme(legend.position = "none")

p4<-ggplot(datos4, aes(nombre,valor,fill=valor,group=valor)) + geom_col(width = 1) +theme(axis.text.x = element_text(size = 7,face="bold",angle = 90, vjust = 0.5, hjust=1))+scale_y_continuous(limits=c(0, 1))+scale_fill_gradient(high = munsell::mnsl("5P 2/12"), low = munsell::mnsl("5P 7/12"))+ggtitle("Cluster 4 Profile")+xlab(NULL)+ylab(NULL)+theme(legend.position = "none")

p5<-ggplot(datos5, aes(nombre,valor,fill=valor,group=valor)) + geom_col(width = 1) +theme(axis.text.x = element_text(size = 7,face="bold",angle = 90, vjust = 0.5, hjust=1))+scale_y_continuous(limits=c(0, 1))+scale_fill_gradient(high = munsell::mnsl("5P 2/12"), low = munsell::mnsl("5P 7/12"))+ggtitle("Cluster 5 Profile")+xlab(NULL)+ylab(NULL)+theme(legend.position = "none")

p6<-ggplot(datos6, aes(nombre,valor,fill=valor,group=valor)) + geom_col(width = 1) +theme(axis.text.x = element_text(size = 7,face="bold",angle = 90, vjust = 0.5, hjust=1))+scale_y_continuous(limits=c(0, 1))+scale_fill_gradient(high = munsell::mnsl("5P 2/12"), low = munsell::mnsl("5P 7/12"))+ggtitle("Cluster 6 Profile")+xlab(NULL)+ylab(NULL)+theme(legend.position = "none")

p7<-ggplot(datos7, aes(nombre,valor,fill=valor,group=valor)) + geom_col(width = 1) +theme(axis.text.x = element_text(size = 7,face="bold",angle = 90, vjust = 0.5, hjust=1))+scale_y_continuous(limits=c(0, 1))+scale_fill_gradient(high = munsell::mnsl("5P 2/12"), low = munsell::mnsl("5P 7/12"))+ggtitle("Cluster 7 Profile")+xlab(NULL)+ylab(NULL)+theme(legend.position = "none")

p8<-ggplot(datos8, aes(nombre,valor,fill=valor,group=valor)) + geom_col(width = 1) +theme(axis.text.x = element_text(size = 7,face="bold",angle = 90, vjust = 0.5, hjust=1))+scale_y_continuous(limits=c(0, 1))+scale_fill_gradient(high = munsell::mnsl("5P 2/12"), low = munsell::mnsl("5P 7/12"))+ggtitle("Cluster 8 Profile")+xlab(NULL)+ylab(NULL)+theme(legend.position = "none")
ggarrange(p1,p2,p3,p4,p5,p6,p7,p8, ncol = 2)
```
